---
title: "Proteomic analysis SMEG"
author: "Marie Locard-Paulet"
date: '`r date()`'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = F)
```

```{r, message=F}
require(ggplot2)
require(knitr)

#source("https://bioconductor.org/biocLite.R")
#biocLite(c("MSnbase", "pRoloc", "pRolocdata", "pRolocGUI"))
require(MSnbase)
require(pRoloc)

source("RFunctions/pdfMLP.R")
```

I load the data before stat (technical repeats aggregated and missing values replaced).

```{r}
load("Data/Data03.Rdata")
```

I remove the rows with only missing values in the quan.

```{r}
mat <- as.matrix(export[,527:542])
naval <- sapply(seq_len(nrow(mat)), function(x) {
  length(mat[x,][is.na(mat[x,])])
})
export <- export[naval == 0,]
```

I create 4 independent MSnet objects with the mean of the log2-transformed normalised values of the technical repeats.

```{r}
# The order of the columns is important when creating the MSnet object:
coln1 <- c(1,5,9,13)+494
coln2 <- coln1+1
coln3 <- coln1+2
coln4 <- coln1+3

coln <- c(coln1, coln2, coln3, coln4)
# The name of the columns has to be the condition . the replicate:
names(export)[coln] <- gsub("[1234]_MeanTech", "", names(export)[coln])
names(export)[coln] <- paste0(names(export)[coln], ".", rep(c(1,2,3,4), each = 4))

hl1 <- readMSnSet2(export, ecol = coln1, fnames = 1)
head(exprs(hl1))
hl2 <- readMSnSet2(export, ecol = coln2, fnames = 1)
head(exprs(hl2))
hl3 <- readMSnSet2(export, ecol = coln3, fnames = 1)
head(exprs(hl3))
hl4 <- readMSnSet2(export, ecol = coln4, fnames = 1)
head(exprs(hl4))
```

I impute the missing values with a noise drawn around the 5\% quantile and a standard deviation estimated from the technical variation of the proteins in the 10\% less intense.

```{r}
lhl <- list(hl1, hl2, hl3, hl4)

# Determining the standard deviation and noise for imputing the missing values:
lstdev <- lapply(lhl, function(x) {
  sapply(1:nrow(as.matrix(exprs(x))), function(y) {
    sd(as.matrix(exprs(x))[y,])
  })
})
matdev <- cbind(lstdev[[1]], lstdev[[2]], lstdev[[3]], lstdev[[4]])

matmean <- export[,grepl("_Mean", names(export))]
matmean <- apply(as.matrix(matmean), 2, as.numeric)
# Remove replacement of missing values in the stat table:
mv <- as.numeric(names(sort(table(matmean), decreasing = T)[1:4]))
matmean[matmean %in% mv] <- NA
valmean <- rowMeans(matmean, na.rm = T)
cutoff <- quantile(valmean, 0.1)
matdev <- matdev[valmean <= cutoff,]
meddev <- median(as.matrix(matdev), na.rm = T)

matval <- cbind(as.matrix(exprs(lhl[[1]])), as.matrix(exprs(lhl[[2]])), as.matrix(exprs(lhl[[3]])), as.matrix(exprs(lhl[[4]])))
noise <- quantile(matval[!is.na(matval)], 0.01)
stdev <- meddev

#lp <- vector(mode = "list", length(lhl))
for (i in seq_along(lhl)) {
  l <- as.matrix(exprs(lhl[[i]]))
  image2(l, main = paste0("Replicate ", i, " before normalisation"))
  # Impute missing value:
  vrep <- runif(length(l),(noise-stdev),noise+stdev)
  l[is.na(l)] <- vrep[is.na(l)]
  exprs(lhl[[i]]) <- l
  l <- lhl[[i]]
  # Normalisation:
  l <- normalise(l, method = "sum")
  
  image2(l, 
        main = paste0("Replicate ", i, ": Values after normalisation and imputation of missing values"))
  plot2D(l, fcol = NULL, col = "black")
  plot2D(l, method = "hexbin")
  lhl[[i]] <- l
}

par(mfrow = c(2,2))
pdf("Figures/LOPIT01.pdf", useDingbats=FALSE, 11.69, 8.27)
for (i in seq_along(lhl)) {
  plot2D(lhl[[i]], method = "hexbin")
}
dev.off()

```

Combination of the data:

```{r}
# Combine the data:
hl <- combine(lhl[[1]], lhl[[2]], lhl[[3]], lhl[[4]])
```


I annotate the data:

```{r}
# Annotate the data:
#fvarLabels(hl)
tab <- read.table("RAW/Quantified proteins 290917_Mapping.txt", sep = "\t", header = T)
tab[,1] <- gsub("_[123456]", "_MeanTech", tab[,1])
tab <- tab[,-4]
tab <- tab[!duplicated(tab),]
names(tab)[3] <- "Replicate"
names(tab)[2] <- "Gradient.Fraction"
tab <- tab[order(tab$Replicate),]
tab$Sample.name <- gsub("[1234]_MeanTech", "", tab$Sample.name)
tab$Sample.name <- paste0(tab$Sample.name, ".", rep(c(1,2,3,4), each = 4))
row.names(tab) <- tab$Sample.name # This is essential.
tab <- tab[,c(3,1,2)]
pData(hl) <- tab
```

```{r}
par(mfrow = c(1, 2), ## creates a two-panel figure
    las = 2, ## axis labels orientation
    cex.axis = .7) ## axis label size
o <- order(hl$Sample.name)
plotDist(hl[, o], pcol = "#00000010", xlab = "")
lines(colMeans(exprs(hl[, o])), col = "red", type = "b")
boxplot(exprs(hl[, o]))

plot2D(hl, fcol = NULL, col = "black")
plot2D(hl, method = "hexbin")

par(mfrow = c(1,1))
pdf("Figures/LOPIT01_combined.pdf", useDingbats=FALSE, 11.69, 8.27)
plot2D(hl, method = "hexbin")
dev.off()
```

# Add markers


<!-- I extract from the stat table the proteins that are specific of a compartment and label them on the plot. -->

```{r, eval = F}
matval <- export[,substr(x = names(export), 1, 2) == "In"]
vec <- rep(NA, nrow(matval))
rs <- rowSums(matval)
for (i in seq_len(nrow(matval))) {
  vec[matval[,1]==1 & rs == 1] <- "Cytoplasm"
  vec[matval[,2]==1 & rs == 1] <- "Mycomembrane"
  vec[matval[,3]==1 & rs == 1] <- "PlasmaMembrane"
  vec[matval[,4]==1 & rs == 1] <- "Secretome"
}
names(vec) <- export[,1]
vec <- vec[!is.na(vec)]

markers <- vec

markersPresAbs <- markers
markersPresAbs <- paste0(markersPresAbs, "-FromDataSet")
names(markersPresAbs) <- names(markers)


hl <- addMarkers(hl, markers)
par(mfrow = c(1, 2))
plot2D(hl, main = "pRolocmarkers from present/absent proteins")
addLegend(hl, cex = .6)
plot2D(hl, dims = c(1, 3), main = "PC1 Vs PC3")
#plot3D(hl)

par(xpd=TRUE, mar = c(4,4,4,16), mfrow = c(1,1))
hlo <- hl[, order(hl$Sample.name)]
col <- c("orange", "red", "forestgreen", "blue")
plotDist(hlo[fData(hlo)$markers == unique(markers)[1],], pcol = col[1], fractions = "Sample.name")
title(main = "Marker occupancy profiles along the gradient")
for (i in 1:length(unique(markers))) {
  matlines(t(exprs(hlo[fData(hlo)$markers == unique(markers)[i], ])), lty = 1, col = col[i], type = "l")
}
legend(17, 0.3, unique(markers), lty = 1, col = col, bty = "n")
```


## Use markers specific for cellular compartments 


I use the table from Laura, it is in the LOPIT folder. These markers have been determined from the literature.

```{r}
fData(hl)$markers <- NULL

markertab <- read.table("LOPIT/protéines marqueurs LOPIT.txt", header = T, stringsAsFactors = F, sep = "\t")
markers <- markertab$localisation
markers[markers == "MycoMembrane - CW"] <- "Mycomembrane"
markers[markers == "Plasma Mb - MP"] <- "PlasmaMembrane"
markers[markers == "Cytosol.Periplasm - CP"] <- "Cytoplasm"
markers[markers == "Secreted Protein - CF"] <- "Secretome"
names(markers) <- markertab$accession

# markers2 <- c(markers, markersPresAbs)
# markers2 <- markers2[!duplicated(names(markers2))]

# hl2 <- hl
hl <- addMarkers(hl, markers) # Only markers from the litterature (for the classification alg.)

# Plot exports:
col <- getStockcol()[1:4]

pdf("Figures/LOPIT02.pdf", useDingbats=FALSE, 8.27, 8.27)
par(mfrow = c(2, 2))
plot2D(hl[, hl$Replicate == 1], main = "Replicate 1", col = col)
legend(-4, -2.6, unique(markers), lty = 1, col = col, bty = "n")
plot2D(hl[, hl$Replicate == 2], main = "Replicate 2", mirrorX = TRUE, col = col)
plot2D(hl[, hl$Replicate == 3], main = "Replicate 3", mirrorX = TRUE, col = col)
plot2D(hl[, hl$Replicate == 3], main = "Replicate 4", mirrorX = TRUE, col = col)
plot2D(hl[, hl$Replicate == 1], dims = c(1, 3), main = "Replicate 1", col = col)
legend(1.6, -1.6, unique(markers), lty = 1, col = col, bty = "n")
plot2D(hl[, hl$Replicate == 2], dims = c(1, 3), main = "Replicate 2", mirrorX = TRUE, col = col)
plot2D(hl[, hl$Replicate == 3], dims = c(1, 3), main = "Replicate 3", mirrorX = TRUE, col = col)
plot2D(hl[, hl$Replicate == 3], dims = c(1, 3), main = "Replicate 4", mirrorX = TRUE, col = col)
dev.off()

pdf("Figures/LOPIT02_combined.pdf", useDingbats=FALSE, 11.69, 6.27)
par(mfrow = c(1, 2))
plot2D(hl, main = "Known markers", col = col, identify = T)
legend(-8, -4, unique(markers), lty = 1, col = col, bty = "n")
plot2D(hl, dims = c(1, 3), main = "Known markers", col = col)
dev.off()
```

# Find protein sub-cellular localisation

```{r}
# generating weithts for classification:
w <- table(getMarkers(hl, verbose = TRUE))
w <- 1/w[names(w) != "unknown"]
```

```{r, eval = TRUE}
# Optimisation of the classifier parameters:
params <- svmOptimisation(hl, fcol = "markers",
                          times = 150, xval = 5,
                          class.weights = w) # I have to check if it is necessary to give weights. Or if I should not give as weights the numbers of proteins in only one compartment in our data set.

save(params, file = "Data/LOPITParams.Rdata")
```

```{r}
load(file = "Data/LOPITParams.Rdata")
plot(params)
levelPlot(params)

(best <- getParams(params))


hl <- svmClassification(hl, params, class.weights = w, fcol = "markers")

par(mar = c(8, 4, 6, 1))
boxplot(fData(hl)$svm.scores ~ fData(hl)$svm, main = "class-specific classification distribution\nlower the better")
```

I choose a threshold corresponding to 33% of the classification scores.

```{r}
#hl <- svmClassification(hl, cost = 16, sigma = 0.1, class.weights = w, fcol = "markers")
(ts <- orgQuants(hl, fcol = "svm", scol = "svm.scores", mcol = "markers", t = .33))
## set point size of each protein to be inversely proportional to the
# ptsze <- exp(fData(hl)$svm.scores) - 1
# ## plot new predictions
# par(mfrow = c(1,2))
# plot2D(hl, fcol = "svm", cex = ptsze, alpha = 0.6)
# addLegend(hl, fcol = "svm", where = "bottomleft", bty = "n", cex = .5)
# plot2D(hl, dims = c(1, 3), fcol = "svm", cex = ptsze)
# 
# plot3D(hl, fcol = "svm", cex = ptsze)



hl <- getPredictions(hl, fcol = "svm", scol = "svm.scores", mcol = "markers", t = ts)
## First remove the markers
preds <- unknownMSnSet(hl)
## Plot a boxplot of the scores of each organelle
par(oma = c(10.5, 0, 0, 0)) ## sets outer margins
boxplot(svm.scores ~ svm, data = fData(preds),
        ylab = "SVM scores", las = 2)



ptsze <- exp(fData(hl)$svm.scores) - 1
par(mfrow = c(1,2))
plot2D(hl, fcol = "svm.pred", cex = ptsze, alpha = 0.6)
addLegend(hl, fcol = "svm", where = "bottomleft", bty = "n", cex = .5)
title("LOPIT prediction results: dim. 1 and 2")
plot2D(hl, dims = c(1, 3), fcol = "svm.pred", cex = ptsze, alpha = 0.6)
title("LOPIT prediction results: dim. 1 and 3")
```

```{r}
sessionInfo()
```
