---
title: "Proteomic analysis SMEG"
author: "Marie Locard-Paulet"
date: '`r date()`'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = F)
```

```{r, message=F}
require(ggplot2)
require(corrplot)
require(gplots)
require(knitr)
require(reshape2)
require(VIM) # visualisation missing data
```

```{r}
# Create colour function for the heatmap:
colvec <- c("darkgreen", "yellow", "red")
pal <- colorRampPalette(colvec)

mlpplot90 <- function(plot) {
        print(plot + theme_minimal() + scale_fill_brewer(palette = "Set1") + theme(axis.text.x = element_text(angle = 90, hjust = 1, size = 8)))
}
```

```{r, echo=T}
load("Data/Data00.Rdata")
```


# Data normalisation 

I remove the proteins with 1 or less unique peptide:

```{r}
mat[input$specific_peptides<=1,] <- NA
```

1/ Log2-transform the matrix

```{r}
matlog <- log2(mat)
hist(matlog)
```

2/ Substract the median of medians

```{r}
med0 <- sapply(1:ncol(matlog), function(x) {median(matlog[,x], na.rm = T)})
med1 <- median(med0)
med <- med0-med1
```

```{r}
# Make export table:
colnames(mat) <- paste0(colnames(mat), "_woMS2")
export <- cbind(input, mat)
colnames(matlog) <- paste0(colnames(matlog), "_Log2")
export <- cbind(export, matlog)

mat <- matlog
```

```{r}
boxplot(matlog, las = 2, main = "Log2-transformed intensities", cex = 0.4, cex.axis = 0.5)
rm(matlog)
for (i in 1:ncol(mat)) {
  mat[,i] <- mat[,i]-med[i]
}
boxplot(mat, las = 2, main = "Log2-transformed intensities after normalisation", cex = 0.4, cex.axis = 0.5)

# Make export table:
colnames(mat) <- sub("_Log2", "_Normalised", colnames(mat))
export <- cbind(export, mat)
```

# Agregation of technical data

I calculate the mean of the technical repeats.

```{r}
colnames(mat) <- sub("_[1234]_Normalised", "", colnames(mat))
vec <- sort(unique(colnames(mat)))
mat2 <- matrix(ncol = length(vec), nrow = nrow(mat))
colnames(mat2) <- vec
rownames(mat2) <- rownames(mat)
matcv <- mat2
for (i in 1:length(vec)) {
  val <- vec[i]
  mattemp <- mat[,colnames(mat)==val]
  m <- rowMeans(mattemp, na.rm = T)
  s <- sapply(1:nrow(mattemp), function(x) {sd(mattemp[x,], na.rm = T)})
  s <- s/m
  mat2[,i] <- m
  matcv[,i] <- s
}
boxplot(mat2, main = "Mean of technical repeats", cex = 0.4, cex.axis = 0.6)
hist(matcv, col = "darkblue", main = "CV of all technical values")
```

```{r}
# Make export table:
colnames(mat2) <- paste0(colnames(mat2), "_MeanTech")
colnames(matcv) <- paste0(colnames(matcv), "_CVTech")
export <- cbind(export, mat2, matcv)

mat <- mat2
rm(list = c("mat2", "mattemp", "matcv"))
```

# Replacement of missing values

Missing values are in red, higher intensities are darker:
```{r}
matrixplot(mat)
```

Keep track and remove the proteins with only missing values:
```{r}
k <- sapply(1:nrow(mat), function(x) length(mat[x,][is.na(mat[x,])]))
ProtMV <- rownames(mat)[k==ncol(mat)]
```

It concerns `r length(ProtMV)`.


I replace ce missing values with the 1\% quantile per biological replicate.

```{r}
gtab <- melt(mat)
g <- ggplot(gtab, aes(x = Var2, y = value, fill = substr(Var2, 1, 2))) + geom_violin()
mlpplot90(g)

# Replacement missing values:
matrep <- mat
repval <- sapply(1:ncol(mat), function(x) {quantile(mat[,x][!is.na(mat[,x])], probs = 0.01)})
for (i in 1:ncol(matrep)) {
  matrep[,i][is.na(matrep[,i])] <- repval[i]
}

# Put back NAs in the proteins from the ProtMV list:
matrep[rownames(matrep) %in% ProtMV,] <- NA

gtab <- melt(matrep)
g <- ggplot(gtab, aes(x = Var2, y = value, fill = substr(Var2, 1, 2))) + geom_violin()
mlpplot90(g)
```

```{r}
# Make export table:
colnames(matrep) <- sub("_MeanTech", "_woMissingVal", colnames(matrep))
export <- cbind(export, matrep)

mat <- matrep
rm(matrep)
```

Visualisation of multivariate analysis:
I center the data.

```{r}
boxplot(mat, las = 2, main = "Log2-transformed data after replacement", cex = 0.4, cex.axis = 0.5)
heatmap.2(scale(mat[complete.cases(mat),], center = T, scale = F), labRow = FALSE, ColSideColors = matcol[match(substr(colnames(mat), 1, 2), matcol[,2]),1], col = pal(300), trace = "none", margins = c(9,7))
```

```{r}
pca <- prcomp(mat[complete.cases(mat),], center = T, scale = F)

tab <- summary(pca)
kable(tab$importance[,1:3])

gtab <- pca$rotation
gtab <- as.data.frame(gtab)
gtab$Condition <- mapping$cellular.compartment[match(sub("_woMissingVal", "", colnames(mat)), sub("_[1234]", "", mapping$Sample.name))]
gtab$Replicate <- mapping$biological.replicate[match(sub("_woMissingVal", "", colnames(mat)), sub("_[1234]", "", mapping$Sample.name))]
gtab$Replicate <- as.character(gtab$Replicate)

ggplot(gtab, aes(x = PC1, y = PC2, col = Condition, label = Replicate, shape = Replicate)) + theme_bw() + geom_point() + scale_colour_manual(values = matcol[match(matcol[,3],levels(gtab$Condition)),1])
```

The data is exported as "Output/ExportedTableBeforeStat.txt"
```{r}
write.table(export, row.names = F, "Output/ExportedTableBeforeStat.txt", sep = "\t")
save(list = c("export", "mapping", "mat", "matcol"), file = "Data/Data01.Rdata")
```






*******************************************************************************
```{r}
sessionInfo()
```
