---
title: "Proteomic analysis SMEG"
author: "Marie Locard-Paulet"
date: '`r date()`'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = F)
```

```{r, message=F}
require(ggplot2)
require(corrplot)
require(gplots)
require(knitr)
```

```{r}
# Create colour function for the heatmap:

colvec <- c("darkgreen", "yellow", "red")
pal <- colorRampPalette(colvec)

```

# Data input

Input: "RAW/Quantified proteins 290917.txt"
I have replaced the "#", "'" and tab in the excel file (Proline export) before saving in tab-delimited format.

**There are rows with no values in the abundance.**

```{r, echo = T}
input <- read.table("RAW/Quantified proteins 290917.txt", header = T, sep = "\t")
mapping <- read.table("RAW/Quantified proteins 290917_Mapping.txt", header = T, sep = "\t")
mapping$Label <- paste0(mapping$cellular.compartment, "-", mapping$biological.replicate, "-", mapping$technical.replicate)
kable(mapping)
```

I keep the columns "abundance" from the Proline table.

```{r}
mat <- input[,c(2,which(grepl("abundance", names(input))))]
mat <- mat[,which(!grepl("raw_", names(mat)))]
na <- mat[,1]
mat <- as.matrix(mat[,2:ncol(mat)])
row.names(mat) <- na
colnames(mat) <- sub("abundance_Laura.", "", colnames(mat))

# Create colour matrix for the figures:
colgroups <- c("brown3", "deepskyblue", "tomato", "slateblue3")
matcol <- cbind(colgroups, unique(substr(colnames(mat), 1, 2)), as.character(mapping$cellular.compartment[match(unique(substr(colnames(mat), 1, 2)), substr(mapping$Sample.name, 1, 2))]))
```

# Data cleaning

I remove the abundance values when no "psm_count":
```{r, eval=T}
# matrix with the number of ms/ms per prot per sample:
matpsm <- input[,c(2,which(grepl("psm_count", names(input))))]
matpsm <- as.matrix(matpsm[,2:ncol(matpsm)])
# Remove abundance values for cells with psm count == 0 or NA:

histinfo <- hist(log(mat), main = "Log intensities, in red the ones removed")
hist(log(mat[matpsm==0 | is.na(matpsm)]), add = T, col = "red")
l <- length(mat[matpsm==0 | is.na(matpsm)])
mat[matpsm==0 | is.na(matpsm)] <- NA
hist(log(mat), main = "Log intensities from identified ions only", col = "steelblue", xlim = c(min(histinfo$breaks),max(histinfo$breaks)))
```

We removed `r l` abundance values with NAs or 0 psm_count.


```{r, eval = F}
image(log(mat))
```

# Quality control


```{r, eval = F}
# Pairwise plots:
pairs(log(mat[,1:16]), pch = ".", main = "plasma membrane")
pairs(log(mat[,17:32]), pch = ".", main = "secretome")
pairs(log(mat[,33:48]), pch = ".", main = "mycomembrane")
pairs(log(mat[,49:64]), pch = ".", main = "cytoplasm")
```

```{r}
boxplot(log(mat), las = 2, cex = 0.4, cex.axis = 0.5, main = "log-transformed abundance across runs")
```

*There is a bias in the normalisation because it was performed in Proline with the data from match between run. More specifically, the median of the "CF" condition is higher than the other cellular compartments.*

## Correlation analysis on the data with no missing values


```{r}
M <- cor(log(mat[complete.cases(mat),]))
corrplot(M, method="color", tl.col = "black", tl.cex = 0.5, mar = c(1,1,1,1), order = "hclust")
par(mar = c(8,2,4,1))

# Heatmap:
heatmap.2(scale(log(mat[complete.cases(mat),]), center = T, scale = F), trace = "none", labRow = FALSE, ColSideColors = matcol[match(substr(colnames(mat), 1, 2), matcol[,2]),1], col = pal(300))
```

## PCA analysis on the data with no missing values

I center the data.

```{r}
pca <- prcomp(log(mat[complete.cases(mat),]), center = T, scale = F)

tab <- summary(pca)
kable(tab$importance[,1:3])

gtab <- pca$rotation
gtab <- as.data.frame(gtab)
gtab$Condition <- mapping$cellular.compartment[match(sub("abundance_Laura.", "", colnames(mat)), mapping$Sample.name)]
gtab$Replicate <- mapping$biological.replicate[match(sub("abundance_Laura.", "", colnames(mat)), mapping$Sample.name)]
gtab$Replicate <- as.character(gtab$Replicate)

ggplot(gtab, aes(x = PC1, y = PC2, col = Condition, label = Replicate, shape = Replicate)) + theme_bw() + geom_point() + scale_colour_manual(values = matcol[match(matcol[,3],levels(gtab$Condition)),1])
```

# Identities of the proteins detected in the cell compartments

Proteins are detected in a compartment if these have an abundance in all the runs.

Venn diagrams:
```{r}
# Make list with one matrix per compartment:
l <- lapply(sort(unique(as.character(mapping$cellular.compartment[match(colnames(mat), mapping$Sample.name)]))), function(x) mat[,as.character(mapping$cellular.compartment[match(colnames(mat), mapping$Sample.name)])==x])
names(l) <- sort(unique(as.character(mapping$cellular.compartment[match(colnames(mat), mapping$Sample.name)])))
l2 <- l
for (i in seq_along(l)) {
  temp <- l2[[i]] 
  temp[temp>0 & !is.na(temp)] <- 1
  l2[[i]] <- temp
}
lsum <- lapply(l2, rowSums, na.rm = T)

lprot <- lapply(1:length(l), function(x) rownames(l[[x]])[lsum[[x]]>=16])
kable(data.frame("Cellular compartment" = names(l), "Number of proteins detected" = sapply(lprot, length)))
#venn(lprot)
venn(list("cytoplasm" = lprot[[1]], "mycomembrane" = lprot[[2]], "plasma membrane" = lprot[[3]], "secretome" = lprot[[4]]))

int <- intersect(lprot[[1]], lprot[[2]])
int <- intersect(lprot[[3]], int)
int <- intersect(lprot[[4]], int)


boxplot(log10(mat[rownames(mat) %in% int,]), las = 2, main = "Proteins common to the 4 compartments", cex = 0.4, cex.axis = 0.5)
heatmap.2(scale(log10(mat[rownames(mat) %in% int,]), center = T, scale = F), labRow = FALSE, ColSideColors = matcol[match(substr(colnames(mat), 1, 2), matcol[,2]),1], col = pal(300), trace = "none")

secprot <- setdiff(lprot[[4]], lprot[[1]])
secprot <- setdiff(secprot, lprot[[2]])
secprot <- setdiff(secprot, lprot[[3]])
write.table(input$description[match(secprot, input$accession)], "Output/secretome.txt", sep = "\t")
secprot <- setdiff(lprot[[2]], lprot[[1]])
secprot <- setdiff(secprot, lprot[[3]])
secprot <- setdiff(secprot, lprot[[4]])
write.table(input$description[match(secprot, input$accession)], "Output/mycomembrane.txt", sep = "\t")
secprot <- setdiff(lprot[[3]], lprot[[1]])
secprot <- setdiff(secprot, lprot[[2]])
secprot <- setdiff(secprot, lprot[[4]])
write.table(input$description[match(secprot, input$accession)], "Output/plasmaMembrane.txt", sep = "\t")
secprot <- setdiff(lprot[[1]], lprot[[3]])
secprot <- setdiff(secprot, lprot[[2]])
secprot <- setdiff(secprot, lprot[[4]])
write.table(input$description[match(secprot, input$accession)], "Output/cytoplasm.txt", sep = "\t")
```

The lists of proteins identified in the compartments are saved in "Output/"

# Data export

I export the data in "Data\"

```{r, echo=T}
save(list = c("input", "mat", "mapping", "matcol"), file = "Data/Data00.Rdata")
```

```{r}
sessionInfo()
```
